### 2015-08-29 psharp.tex
more text

\subsection{Overview}
\label{sec:psharp:overview}

Draft

\subsection{Liveness checking}
\label{sec:psharp:liveness}

Draft

\subsection{Handling intra-machine concurrency}
\label{sec:psharp:async}

Async/Await

custom schedulers, etc

\subsection{Other}
\label{sec:psharp:other}

Logs/traces -> user can extend them?

\PDComment{mention dependency injection pattern?}