### 2015-09-20 paper.tex
updates to abstract

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}

\usepackage{usenix,epsfig,endnotes}

%\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
%\usepackage{microtype}
%\usepackage{flushend}

%\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{pifont}

\usepackage{hyperref}
\usepackage{multirow}
\usepackage{listings}
\usepackage{booktabs}
%\usepackage{tikz}
\usepackage{xspace}

\usepackage{balance}

\newcommand{\psharp}{P\#\xspace}
\newcommand{\csharp}{C\#\xspace}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\usepackage{color}
\definecolor{light-gray}{gray}{0.40}
\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
\newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
\newcommand{\PTComment}[1]{\textcolor{light-gray}{Paul: #1}}

\definecolor{bluekeywords}{rgb}{0.13,0.13,1}
\definecolor{greencomments}{rgb}{0,0.5,0}
\definecolor{redstrings}{rgb}{0.9,0,0}

\lstset{language=[Sharp]C,
showspaces=false,
showtabs=false,
breaklines=true,
showstringspaces=false,
breakatwhitespace=true,
escapeinside={(*@}{@*)},
morekeywords={OnEvent, nameof, var, state, cold, hot},
commentstyle=\color{greencomments},
keywordstyle=\color{bluekeywords}\bfseries,
stringstyle=\color{redstrings},
basicstyle=\ttfamily\scriptsize,
columns=fullflexible
}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)}

%for single author (just remove % characters)
\author{
{\rm Pantazis Deligiannis$^\dagger$, Matt McCutchen$^\diamond$, Paul Thomson$^\dagger$, Shuo Chen$^\star$}\\
{\rm Alastair F. Donaldson$^\dagger$, John Erickson$^\star$, Cheng Huang$^\star$, Akash Lal$^\star$}\\
{\rm Rashmi Mudduluru$^\star$, Shaz Qadeer$^\star$, Wolfram Schulte$^\star$}\\\\
$^\dagger$Imperial College London, $^\diamond$Massachusetts Institute of Technology, $^\star$Microsoft\\
} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
Testing distributed storage systems is challenging due to multiple sources of nondeterminism. Conventional testing techniques, such as unit and stress testing, are ineffective in preventing serious but subtle bugs from reaching production. Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code. In this paper, we present a new methodology for testing production distributed systems. Our approach applies systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.

Our methodology has been applied to three production distributed storage systems in the Microsoft Azure platform. In the process, numerous bugs were identified, reproduced, confirmed and fixed. These bugs required a subtle combination of concurrency and failures, making them extremely difficult to find with conventional testing techniques. An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace, which dramatically increases the productivity of debugging.

\section{Introduction}
\label{sec:intro}

\input{intro}

\section{Motivating Example}
\label{sec:motivation}

\input{motivation}

\section{Our Approach to Testing Distributed Systems}
\label{sec:overview}

\input{overview}

\section{Testing Azure Storage vNext with \psharp}
\label{sec:method}

\input{methodology}

\section{Additional Case Studies}
\label{sec:cases}

\input{experience}

\section{Quantifying the Cost of Using \psharp}
\label{sec:eval}

\input{evaluation}

\section{Related Work}
\label{sec:rw}

\input{related}

\section{Conclusion}
\label{sec:conclusion}

\input{conclusion}

%\section{Acknowledgments}

%Draft.

% Balances the bibliography in the last page
\balance

{\footnotesize \bibliographystyle{acm}
\bibliography{references}}

\end{document}